---
title: "Tema 54 ejemplos"
author: "Gabriel Dávila"
date: "7/4/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Ejemplos Tema 54


```{r, eval = TRUE}
x = seq(3,50,by=3.5)
x
x[3]
x[8]
x[length(x)]
x[length(x)-1]
x[length(x)-2]
x[-3]
x[4:8]
x[8:4]
x[seq(2,length(x), by=2)]
x[seq(1,length(x), by=2)]
x[-seq(2,length(x), by=2)]
x[(length(x)-3):length(x)]
x[c(1,5,6)]
x[x>30]
x[x>20 & x<40]
x[x!=3 & x!=17]
x[x<10|x>40]
x[x>=10]
x[x>10]
x[!x<10]
x[x%%2==0]
x[x%%2==1]
x>30
x = c(1,7,4,2,4,8,9,2,0)
y = c(5,2,-3,-7,-1,4,-2,7,1)
x[y>0]
x
which(x>4)
x[which(x>4)]
x[x>4]
which(x>2 & x<8)
which(x<5 | x%%2==0)
x[which(x<5 | x%%2==0)]
which(x%%2==0)
which.min(x)
which(x == min(x))
which.max(x)
which(x == max(x))
fix(x)
which.max(x)
which(x == max(x))
x[x>4 & x<7]
which(x>4 & x<7)
z = c()
z
a = NULL
a
b = c(z, 2, a, 7)
b
```


## Ejemplos Tema 55. Los valores NA

```{r}
2^pi > pi^2
2^pi
pi^2
12345678%%9==0
12345678/9
length(1:100)>4
length(1:100)>400
x = 1:10
x[3]
x[3] = 32
x
x[11]
x[11] = 15
x
x[2:5] = x[2:5]+3
x
x[(length(x)-2):length(x)] = 0
x
x[length(x)+5]=9
x
sum(x)
prod(x)
cumsum(x)
mean(x)
sum(x)
sum(x, na.rm = TRUE)
mean(x, na.rm = TRUE)
which(x == NA)
is.na(x)
which(is.na(x))
x[which(is.na(x))]
x
y = x
y
y[is.na(y)]
y[is.na(y)] = mean(y, na.rm=TRUE)
y
x[is.na(x)]
x[!is.na(x)]
sum(x[!is.na(x)])
#cumsum(x, na.rm == TRUE)
cumsum(x[!is.na(x)])
# Opción no recomendada, eliminar NAs definitivamente:
na.omit(x)
sum(na.omit(x))
cumsum(na.omit(x))
x_clean = na.omit(x)
x_clean
attr(x_clean, "na.action") = NULL
x_clean
attr(x_clean, "class") = NULL
x_clean
```

## 56. Factores

gender = c("M", "H", "H","M","M","M","M", "H", "H")
gender.fact = factor(gender)
gender.fact2 = as.factor(gender)
gender.fact2
gender.fact3 = factor(gender, levels = c("H","M","NB"))
gender.fact3
gender.fact4 = factor(gender, levels = c("H","M","NB"), labels = c("Hombre","Mujer","No Binario"))
gender.fact4
levels(gender.fact)
levels(gender.fact2)
levels(gender.fact3)
levels(gender.fact4)
levels(gender.fact4) = c("Masculino","Femenino","Fluido")
levels(gender.fact4)
gender.fact4


notas = c(1,4,3,2,3,2,4,3,1,2,3,4,2,3,4)
notas
notas.factor = factor(notas)
notas.factor
levels(notas.factor)
levels(notas.factor) = c("Suspendido", "Suficiente", "Notable", "Excelente")
levels(notas.factor)
notas.factor

notas
levels(notas.factor) = c("Suspendido", "Aprobado", "Aprobado", "Aprobado")
notas.factor



notas
ordered(notas, labels = c("Sus", "Suf", "Not", "Exc"))


## 57. Listas

x = c(1,5,-2,6,-7,8,-3,4,-9)
x
L = list(nombre = "Temperaturas", datos = x, media = mean(x), sumas = cumsum(x))
L$media
L$datos
L$nombre
L[[3]]
L[[2]]

# 58. Matrices

matrix(0, nrow = 3, ncol = 5)

vec = 1:12
matrix(vec, nrow = 3)


x = rbind(matrix(vec, nrow = 3), c(1,2,3,4), c(-1,-2,-3,-4))
x

diag(5, nrow = 3)
diag(c(1,2,3,4))

M = matrix(1:12, nrow= 5,)
M
M[2,3]
M[2,]
M[,3]
M[c(2,3,5), 1:2]

diag(M)
nrow(M)
ncol(M)
dim(M)
sum(M)
prod(M)
mean(M)

colSums(M)
rowSums(M)
colMeans(M)

apply(M, MARGIN = 1, FUN = function(x){sqrt(sum(x^2))})

apply(M, MARGIN = 2, FUN = function(x){sqrt(sum(x^2))})

apply(M, MARGIN = c(1,2), FUN = function(x){x^2})

# 59. Un repaso de álgebra lineal

A = rbind(c(2,0,2),c(1,2,3),c(0,1,3))
A
B = rbind(c(3,2,1),c(1,0,0),c(1,1,1))
B

A*B
B*A

A%*%B
B%*%A

A*A
B*B*B

A%^%2
A^2
B%^%3
B^3
(B%*%B)%*%B


M = rbind(c(1,4,2),c(0,1,3),c(1,8,9))
M

qr(M)$rank

solve(M)

solve(M)%*%M
round(solve(M)%*%M)
solve(M, c(1,2,3))
eigen(M)

eigen(M)$values

eigen(M)$vectors
round(eigen(M)$vectors,2)

M = rbind(c(2,6,-8),c(0,6,-3),c(0,2,1))
M
P = eigen(M)$vectors
P
Pinv = solve(P)
Pinv
D = diag(eigen(M)$values)
D

R = (P%*%D)%*%Pinv
R

M == round(R,3)


C = matrix(c(3-2i, 5+3i, 1+2i, 2-1i), nrow = 2, byrow = T)
C

C%*%C

eigen(C)

solve(C, c(1-1i, 4))
det(C)

detMatComplex = prod(eigen(C)$values)
detMatComplex


